\documentclass[./\jobname.tex]{subfiles}
\begin{document}
\chapter{Conclusion}
This chapter summarises what has been learnt about the performance of the algorithms during the experiments. Further, remarks on advantages and disadvantages of the proposed strategy are briefly stated. 

Using JADE could not replicate the objectively good performance observed with \gls{cma_es} in \cite{chaquet_using_2019} and other related work. The reason for this is not completely apparent. The poor convergence might be due to some bad parameter choices. Otherwise, JADE might not be as well suited for that sort of fitness function as a \gls{cma_es} is. 

To speed up the optimisation process, the parallel JADE is implemented. Thus, the already present parallel hardware structure can be used. Depending on the problem \gls{pde}, evaluating the population in parallel significantly speeds up the solving time by at least a factor of 2.6. The parallel JADE works as expected, however it does not reduce the computational effort. Since nowadays, single-core machines do not exist any more, a parallel structure should be the standard for all newly created software. At the very least, new algorithms should be designed with the possibility for parallelisation in mind. 

The purpose of the adaptive kernels scheme is to reduce the search dimension and increase the number of kernels only when necessary. It works very well on the testbed \gls{pde} 0A. This \gls{pde} fulfils the assumptions that have been made for this algorithm. Mixed results have been observed on all other testbed problems. The experiments suggest that the problems could be overcome with an algorithm that converges faster. 

The present fitness function might not be the perfect formulation of the problem. Ideally, the quality of the approximation should be tightly linked to the fitness value. However, experiments have shown that this might not always be the case. Specifically, on \gls{pde} 5 more \gls{nfe} result in a significantly worse approximation quality. Thus, a new kernel called \gls{gsk} is created. The goal for this kernel is to alter the fitness function and thus enhance the correlation between fitness function and quality. The idea works as intended and the results on that \gls{pde} do get significantly better. Although the \gls{gsk} exhibits all the features of the Gauss kernel, the quality of the results for all other \gls{pde}s are mixed. This might be due to the larger search space that comes with the new kernel. More experiments investigating the connection of the adaptive scheme with the \gls{gsk} could provide more insight into that subject.

Obviously, this \gls{ci} approach to solve \gls{pde}s is not as effective as the \gls{fem} method. \gls{fem} is specifically designed to work on elliptic \gls{pde}s and convergence is proven. However, the \gls{ci} method is a universal approach that should be able to solve all sorts of differential equations. 

Contrary, the ``universal solver'' attribute might also be the greatest disadvantage of this strategy. Currently, there is no guarantee for convergence. The simple implementation and the relatively straight forward approach could misguide the user into applying this strategy, although there might exist better solvers for the problem at hand. 

Since there is no guarantee on the quality of the solution, it is potentially risky to apply the \gls{ci} strategy. The method should work best for problems, where a vague idea on the structure of the solution is already present. A perfect example for such a field of application could be image processing, specifically ``shape from shading'' (\cite{horn_shape_1970}). With this method, the shape of an object can be computed directly from one image. This is done by using the shadows cast by a single light source. At the core of the problem stands a \gls{pde} that needs to be solved. Since the image of the object is available to the user, additional quality checks on the approximation can be performed. Surely, there are other applications that could benefit from this solver strategy. 

\end{document}